Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-02-26 paper.tex
tweak

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{syntax}

@@

\begin{verbatim}
e \in Exp ::= ae
            | (let ([x (f ae)]) e)
ae \in AExp ::= x
              | lam
lam \in Lam ::= (lambda (x) e)
x ... are variable names
\end{verbatim}

In ANF, all the function applications must be administrated in a \texttt{let} expression,
and then bounded to a variable name under current environment.
We can use \texttt{letrec} to create mutually recursive function bindings. But \texttt{letrec}
only accepts atomic expressions that appear at the right-hand side of a binding.
The atomic expressions $ae$ are either a variable, or a literal \texttt{lambda} term, which
both can be evaluated in a single step.

@@

\begin{verbatim}
sealed trait Expr
case class Binding(x: String, e: Expr)

case class Var(x: String) extends Expr
case class App(e1: Expr, e2: Expr) extends Expr
case class Lam(x: String, body: Expr) extends Expr
case class Let(x: String, e: App, body: Expr) extends Expr
\end{verbatim}

@@

\begin{verbatim}
type Addr = Int
type Env = Map[String, Addr]
type Store = Map[Addr, Storable]

@@

case class State(e: Expr, env: Env, store: Store, k: Kont)
\end{verbatim}

It is worth noting that the continuation class \texttt{Kont} is defined as a list of frames.
The frame class \texttt{Frame} stores the information of call-site, i.e., the information that
can be used to resume the interrupted computation.
A \texttt{Frame} constitutes of a variable name \texttt{x} to be bind later, a control expression
that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)

\subsubsection{Single-step Transition}
Before go into describing how the machine evaluates expressions, we firstly define several helper functions.
As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
a literal \texttt{lambda} term. So the helper function \texttt{atomicEval} handles these two
cases and evaluates atomic expressions to closures in a straightforward way.
The \texttt{alloc} function generates a fresh address, and always allocates an unique integer under the domain
of store.
The \texttt{isAtomic} function is used as a predicate of whether the expression is atomic.

\begin{verbatim}
def atomicEval(e: Expr, env: Env, store: Store): Storable = e match {
  case Var(x) => store(env(x))
  case lam@Lam(x, body) => Clos(lam, env)
}
def alloc(store: Store): Addr = store.keys.size + 1
def isAtomic(e: Expr): Boolean = e.isInstanceOf[Var] || e.isInstanceOf[Lam]
\end{verbatim}

Now, we can faithfully describe the state transition function \texttt{step},
that given a machine state, function \texttt{step} determines its successor state.

\begin{verbatim}
def step(s: State): State = s match {
  case State(Let(x, App(f, ae), e), env, store, k) if isAtomic(ae) =>
    val Clos(Lam(v, body), env_c) = atomicEval(f, env, store)
    val addr = alloc(store)
    val new_env = env_c + (v -> addr)
    val new_store = store + (addr -> atomicEval(ae, env, store))
    val frame = Frame(x, e, env)
    State(body, new_env, new_store, frame::k)
  case State(ae, env, store, k) if isAtomic(ae) =>
    val Frame(x, e, env_k)::ks = k
    val addr = alloc(store)
    val new_env = env_k + (x -> addr)
    val new_store = store + (addr -> atomicEval(ae, env, store))
    State(e, new_env, new_store, ks)
}
\end{verbatim}

The first case is that the control of current state is a \texttt{Let} expression,
and its right-hand side is an function application.
By calling function \texttt{atomicEval}, we obtain the closure that callee \texttt{f} stands for.
The successor state's control then transfers to the \texttt{body} expression of closure,
with an updated environment and an update store. The new environment is extended
from closure's environment and mapped \texttt{v} to a fresh address \texttt{addr}.
The new store is extended with \texttt{addr} mapping to the value of \texttt{ae},
which is evaluated from \texttt{atomicEval}.
In the meanwhile, a new frame \texttt{frame} is pushed onto the stack \texttt{k}.
This \texttt{frame} contains the variable name \texttt{x} at left-hand side position of \texttt{Let},
the body expression of \texttt{Let}, and the lexical environment of the body expression.

The second case for \texttt{step} is the control is transferred to an atomic expression.
We extract the top frame of continuations firstly.
The control (that is an atomic expression) of current state is the evaluated term
that being bind to the variable \texttt{x} from the top frame.
The environment and store are updated with \texttt{x} mapping to the closure value of \texttt{ae}.
Then the successor state is transferred to expression \texttt{e} from the top frame,
which is the body of a \texttt{Let} expression, with the updated environment, store, and
the rest of stack \texttt{ks}.

@@

To run the program, we first use function \texttt{inject} to construct an initial machine
state given an expression \texttt{e}. The initial state contains an empty environment, 
store and stack.
\begin{verbatim}
def inject(e: Expr): State = State(e, Map(), Map(), Nil)
\end{verbatim}

Then the \texttt{drive} function is used to evaluate
to a final state by iteratively apply \texttt{step} on current state until we reach a state
that the control is an atomic expression and the continuation stack is empty.
And of course, we can extract the value from the final state at last.

\begin{verbatim}
def drive(s: State): State = s match {
  case State(ae, _, _, Nil) if isAtomic(ae) => s
  case _ => drive(step(s))
}
def eval(e: Expr): State = drive(inject(e))
\end{verbatim}

@@

For clarity, we divide the store into binding store \texttt{BStore} and continuation store \texttt{KStore}. 
The binding store maps binding addresses to sets of closure values, and continuation store maps
continuation addresses to sets of continuations.
We define a generic class \texttt{Store[K,V]} that performs join when updating elements
in a store. By parameterizing \texttt{Store[K,V]} with \texttt{[BAddr, Storable]} and 
\texttt{[KAddr, Cont]}, we obtain \texttt{BStore} and \texttt{KStore}, perspectively.

@@

\begin{verbatim}
case class Store[K,V](map: Map[K, Set[V]]) {
  def apply(addr: K): List[V] = map(addr).toList
  def update(addr: K, d: Set[V]): Store[K,V] = {
    val oldd = map.getOrElse(addr, Set())
    Store[K, V](map ++ Map(addr -> (d ++ oldd)))
  }
  def update(addr: K, sd: V): Store[K,V] = update(addr, Set(sd))
}
type BStore = Store[BAddr, Storable]
type KStore = Store[KAddr, Cont]
\end{verbatim}

The co-domain of binding stores \texttt{Storable} is the same
as we defined for CESK machines.
The co-domain of continuation stores \texttt{Cont} is comprised of 
a frame \texttt{Frame} and a continuation address \texttt{KAddr}.
To mimic the runtime call stack, \texttt{KAddr} plays the role of
representing the remaining stack frames.
TODO: talk about nondet of cont deref?

\begin{verbatim}
abstract class Storable 
case class Clos(v: Lam, env: Env) extends Storable

case class Frame(x: String, e: Expr, env: Env)
case class Cont(frame: Frame, kaddr: KAddr)
\end{verbatim}

As a consequence, the components of a state is also changed: 
the store is divided into binding store and continuation store;
the continuation stack becomes an address.
By dereferencing this address in continuation store, we can retrieve the 
actual transfer of controls. The definition of environment \texttt{Env}
remains the same.

\begin{verbatim}
case class State(e: Expr, env: Env, bstore: BStore, 
                 kstore: KStore, k: KAddr, time: Time)
\end{verbatim}

\subsubsection{Allocating Addresses}
Up to now, we did not describe much about the adresses in stores and the 
time-stamp \texttt{Time}.
But they are key ingredients to achieve analyses with different sensitivities,
as well as a finite state space analysis.
To effectively approximate the runtime behavior, a finite 
program contour \texttt{time} that encodes the program execution history is introduced.
Here, we use a list of execution contexts (expressions) to reprent that, and
as we will see in section \ref{kcfainst}, by using different \texttt{tick}
function on time we may obtain a family of analyses.

\begin{verbatim}
type Time = List[Expr]
\end{verbatim}

As we mentioned previously, the space of addresses is finite in AAM.
Binding addresses are parameterized by variable names and the time that creates this binding,
which are both finite.
Continuation addresses \texttt{KAddr} has two variants: 
1) \texttt{Halt} which corresponds to the empty stack, and
2) \texttt{ContAddr} consists of the entry expression of callee and the time that
creates this address, and they are also finite. 
(TODO: why target expression, allocation polyvariance)

\begin{verbatim}
case class BAddr(x: String, time: Time)

abstract class KAddr
case object Halt extends KAddr
case class ContAddr(tgt: Expr, time: Time) extends KAddr
\end{verbatim}

We use two helper functions \texttt{allocBind} and \texttt{allocKont} that will be used later
to allocate binding addresses and continuation addresses.

\begin{verbatim}
def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
def allocKont(tgtExpr: Expr, time: Time): KAddr = ContAddr(tgtExpr, time)
\end{verbatim}

@@

Since dereferencing an address becomes non-deterministic, our \texttt{atomicEval}
function also truns into non-deterministic. Given an atomic expression \texttt{e},
\texttt{atomicEval} returns a set of storable values, i.e., closures, to the caller.
If the expression is just a \texttt{lambda} term, then the returned set is singleton.

\begin{verbatim}
def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
  case Var(x) => bstore(env(x))
  case lam@Lam(x, body) => Set(Clos(lam, env))
}
\end{verbatim}

The structure of function \texttt{step} is similar with concrete CESK machines, 
except non-determinism which makes \texttt{step} returns a list of reachable 
successor states.
In the beginning of \texttt{step}, we call function \texttt{tick} to update 
the time.
Then we have two cases to consider: 1) Current control expression is \texttt{Let},
and \texttt{App(f, ae)} is being bind to \texttt{x}.
Fristly we retrieve the set of closures that \texttt{f} may represents.
For each closure in the set, we do almost the same as what we did in concrete 
CESK machines. But note that since the continuation is allocated on store (\texttt{kstore}),
a new continuation address \texttt{new_kaddr} is constructed, and a new frame
\texttt{Frame(x, e, env)} paired with the current continuation address \texttt{kaddr}
is joined into this address. Finally, a list of states is generated.

2) An atomic expression \texttt{ae} sits on the control position of the state.
The value of \texttt{ae} is being returned to its caller. 
We dereference the continuation address \texttt{kaddr} and obtain a set of
continuations \texttt{conts}.
For each continuation in the set, we construct an environment based on the
environment of the frame (\texttt{env_f}), and bind \texttt{x} to a newly created binding 
address \texttt{baddr}; also update the store with \texttt{baddr} and 
the values that \texttt{ae} represents. In every generated states,
the control becomes the expression \texttt{e} in the frame,
and as we can tell from the name, continuation address \texttt{f_kaddr} 
also comes from the frame.

\begin{verbatim}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) 
      if isAtomic(ae) =>
      val closures = atomicEval(f, env, bstore).toList
      for (Clos(Lam(v, body), env_c) <- closures) yield {
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
        val new_kaddr = allocKont(body, new_time)
        val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
        State(body, new_env, new_bstore, new_kstore, new_kaddr, new_time)
      }
    case State(ae, env, bstore, kstore, kaddr, time) 
      if isAtomic(ae) =>
      val conts = kstore(kaddr).toList
      for (Cont(Frame(x, e, env_f), f_kaddr) <- conts) yield {
        val baddr = allocBind(x, new_time)
        val new_env = env_f + (x -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(e, new_env, new_store, kstore, f_kaddr, new_time)
      }
  }
}
\end{verbatim}

@@

\begin{verbatim}
def k: Int = 0
def tick(s: State): Time = (s.e::s.time).take(k)
\end{verbatim}

@@

To run (analyze) a program, as CESK machine, we use \texttt{inject} to construct
the initial state given a program. 
Note that the initial continuation store has a special mapping that maps continuation address
\texttt{Halt} to empty set of continuations.
We also provide an empty program contour as initial time.
\begin{verbatim}
def inject(e: Expr): State = 
  State(e, Map(), 
        Store[BAddr, Storable](Map()), 
        Store[KAddr, Cont](Map(Halt -> Set())), 
        Halt, List())
\end{verbatim}

However, comparing with concrete CESK machine, the
\texttt{drive} function performs collecting semantics instead of valuation
semantics. That is, for the purpose of analyzing programs, the function \texttt{drive}
collects all the intermediate machine states as the program is abstractly executing.
The following code shows a variant of work list algorithm that always applies \texttt{step}
function to the first element \texttt{hd} of the work list \texttt{todo} if \texttt{hd}
is unseen, then adds the result of \texttt{step} to the rest of work list and in 
the meanwhile adds \texttt{hd} to the explored states set.
If the work list is empty, the function just returns the set of reachable states up to now.

\begin{verbatim}
def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
  case Nil => seen
  case hd::tl if seen.contains(hd) => drive(tl, seen)
  case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
}

def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
\end{verbatim}

Finally, a user could use function \texttt{analyze} to obtain all reachable states 
for a given program.

@@

\begin{verbatim}
(let ([id (lambda (z) z)])
  (let ([x (id 1)])
    (let ([y (id 2)])
      x)))
\end{verbatim}

In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
the return-flow of invocation \texttt{(id 2)} 
is going to both call-sites \texttt{(id 1)} and \texttt{(id 2)}. Therefore, the
returned value \texttt{2} for variable \texttt{y} is also propagated to 
variable \texttt{x} and then imprecise analysis result is arose.
This return-flow merging is inevitable even we increase the context-sensitivity.
If we use the monovariant analysis 0CFA, the analyzing result would be \texttt{x} and 
\texttt{y} point to set \texttt{\{1, 2\}}, because it not distinguishes
\texttt{z} at different call-sites.
Under $1$-CFA, the algorithm is able to distinguish that variable \texttt{z} of function
\texttt{id} has two different values at two call-sites, so variable \texttt{y} 
would not be polluted by \texttt{1}, but it is still helplessness for distinguish
return-flows.
TODO: do need more details to explain this example? (as in P4F paper)

@@

In the definition of state \texttt{State}, continuation store disappears, and
\texttt{konts} becomes a list of frames. Other components keep the same.

\begin{verbatim}
case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
\end{verbatim}

The state transition function \texttt{step} is still non-determinism, but
the only non-determinism happens when dereferencing the callee \texttt{f} 
from the store at call-site \texttt{App(f, ae)}. 
For each closure of \texttt{f} in the set, we do exactly
the same as what we did in concrete CESK machines: a new frame is constructed
and pushed onto the stack.
In the second case, since the continuation \texttt{konts} is a list of 
frames, we can simply decompose it to differentiate an empty stack
or a frame exists on the top of stack. 
If the stack is empty, then empty successors is returned, because the computation
is finished.
Otherwise, we matched a frame \texttt{Frame(x, e, env_f)}. 
This transition is deterministic, because we only have one frame, this frame
is then popped and the new state uses the rest of stack.

\begin{verbatim}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
        val frame = Frame(x, e, env)
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(body, new_env, new_store, frame::konts, new_time)
      }
    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
      konts match {
        case Nil => List()
        case Frame(x, e, env_f)::konts =>
          val baddr = allocBind(x, new_time)
          val new_env = env_f + (x -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          List(State(e, new_env, new_store, konts, new_time))
      }
  }
}
\end{verbatim}

@@

\begin{verbatim}
(letrec ([f1 (lambda (x) 
               (let ([x1 (f2 x)]) x1))]
         [f2 (lambda (y)
               (let ([y1 (f1 y)]) y1))])
  (let ([z (f1 1)])
    z))
\end{verbatim}

\texttt{f1} and \texttt{f2} call each others, and the stack will be interleaved
pushing \texttt{f1} and \texttt{f2} onto the top of current stack, but not existing 
two stack components are the same in the state space.

@@

\begin{verbatim}
case class P4FContAddr(tgt: Expr, tgtEnv: Env, time: Time) extends KAddr
def allocKont(tgtExpr: Expr, tgtEnv: Env, time: Time): KAddr = 
  P4FContAddr(tgtExpr, tgtEnv, time)
\end{verbatim}

@@

>>>>

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.9]{beramono}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor,colortbl}
\usepackage{url}
\usepackage{listings}
\usepackage{paralist}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{flushend}

\input{macros}

\lstMakeShortInline[keywordstyle=,%
              flexiblecolumns=false,%
              %basewidth={0.56em, 0.52em},%
              mathescape=false,%
              basicstyle=\tt]@

@@

\begin{lstlisting}
e $\in$ Exp ::= ae
          | (let ([x (f ae)]) e)
ae $\in$ AExp ::= x            
            | lam
lam $\in$ Lam ::= (lambda (x) e)
x ... are variable names
\end{lstlisting}

In ANF, all the function applications must be administrated in a @let@ expression,
and then bounded to a variable name under current environment.
We can use @letrec@ to create mutually recursive function bindings. But @letrec@
only accepts atomic expressions that appear at the right-hand side of a binding.
The atomic expressions $ae$ are either a variable, or a literal @lambda@ term, which
both can be evaluated in a single step.

@@

\begin{lstlisting}
sealed trait Expr
case class Binding(x: String, e: Expr)

case class Var(x: String) extends Expr
case class App(e1: Expr, e2: Expr) extends Expr
case class Lam(x: String, body: Expr) extends Expr
case class Let(x: String, e: App, body: Expr) extends Expr
\end{lstlisting}

@@

\begin{lstlisting}
type Addr = Int
type Env = Map[String, Addr]
type Store = Map[Addr, Storable]

@@

case class State(e: Expr, env: Env, store: Store, k: Kont)
\end{lstlisting}

It is worth noting that the continuation class @Kont@ is defined as a list of frames.
The frame class @Frame@ stores the information of call-site, i.e., the information that
can be used to resume the interrupted computation.
A @Frame@ constitutes of a variable name @x@ to be bind later, a control expression
that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)

\subsubsection{Single-step Transition}
Before go into describing how the machine evaluates expressions, we firstly define several helper functions.
As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
a literal @lambda@ term. So the helper function @atomicEval@ handles these two
cases and evaluates atomic expressions to closures in a straightforward way.
The @alloc@ function generates a fresh address, and always allocates an unique integer under the domain
of store.
The @isAtomic@ function is used as a predicate of whether the expression is atomic.

\begin{lstlisting}
def atomicEval(e: Expr, env: Env, store: Store): Storable = e match {
  case Var(x) => store(env(x))
  case lam@Lam(x, body) => Clos(lam, env)
}
def alloc(store: Store): Addr = store.keys.size + 1
def isAtomic(e: Expr): Boolean = e.isInstanceOf[Var] || e.isInstanceOf[Lam]
\end{lstlisting}

Now, we can faithfully describe the state transition function @step@,
that given a machine state, function @step@ determines its successor state.

\begin{lstlisting}
def step(s: State): State = s match {
  case State(Let(x, App(f, ae), e), env, store, k) if isAtomic(ae) =>
    val Clos(Lam(v, body), env_c) = atomicEval(f, env, store)
    val addr = alloc(store)
    val new_env = env_c + (v -> addr)
    val new_store = store + (addr -> atomicEval(ae, env, store))
    val frame = Frame(x, e, env)
    State(body, new_env, new_store, frame::k)
  case State(ae, env, store, k) if isAtomic(ae) =>
    val Frame(x, e, env_k)::ks = k
    val addr = alloc(store)
    val new_env = env_k + (x -> addr)
    val new_store = store + (addr -> atomicEval(ae, env, store))
    State(e, new_env, new_store, ks)
}
\end{lstlisting}

The first case is that the control of current state is a @Let@ expression,
and its right-hand side is an function application.
By calling function @atomicEval@, we obtain the closure that callee @f@ stands for.
The successor state's control then transfers to the @body@ expression of closure,
with an updated environment and an update store. The new environment is extended
from closure's environment and mapped @v@ to a fresh address @addr@.
The new store is extended with @addr@ mapping to the value of @ae@,
which is evaluated from @atomicEval@.
In the meanwhile, a new frame @frame@ is pushed onto the stack @k@.
This @frame@ contains the variable name @x@ at left-hand side position of @Let@,
the body expression of @Let@, and the lexical environment of the body expression.

The second case for @step@ is the control is transferred to an atomic expression.
We extract the top frame of continuations firstly.
The control (that is an atomic expression) of current state is the evaluated term
that being bind to the variable @x@ from the top frame.
The environment and store are updated with @x@ mapping to the closure value of @ae@.
Then the successor state is transferred to expression @e@ from the top frame,
which is the body of a @Let@ expression, with the updated environment, store, and
the rest of stack @ks@.

@@

To run the program, we first use function @inject@ to construct an initial machine
state given an expression @e@. The initial state contains an empty environment, 
store and stack.
\begin{lstlisting}
def inject(e: Expr): State = State(e, Map(), Map(), Nil)
\end{lstlisting}

Then the @drive@ function is used to evaluate
to a final state by iteratively apply @step@ on current state until we reach a state
that the control is an atomic expression and the continuation stack is empty.
And of course, we can extract the value from the final state at last.

\begin{lstlisting}
def drive(s: State): State = s match {
  case State(ae, _, _, Nil) if isAtomic(ae) => s
  case _ => drive(step(s))
}
def eval(e: Expr): State = drive(inject(e))
\end{lstlisting}

@@

For clarity, we divide the store into binding store @BStore@ and continuation store @KStore@. 
The binding store maps binding addresses to sets of closure values, and continuation store maps
continuation addresses to sets of continuations.
We define a generic class @Store[K,V]@ that performs join when updating elements
in a store. By parameterizing @Store[K,V]@ with @[BAddr, Storable]@ and 
@[KAddr, Cont]@, we obtain @BStore@ and @KStore@, perspectively.

@@

\begin{lstlisting}
case class Store[K,V](map: Map[K, Set[V]]) {
  def apply(addr: K): List[V] = map(addr).toList
  def update(addr: K, d: Set[V]): Store[K,V] = {
    val oldd = map.getOrElse(addr, Set())
    Store[K, V](map ++ Map(addr -> (d ++ oldd)))
  }
  def update(addr: K, sd: V): Store[K,V] = update(addr, Set(sd))
}
type BStore = Store[BAddr, Storable]
type KStore = Store[KAddr, Cont]
\end{lstlisting}

The co-domain of binding stores @Storable@ is the same
as we defined for CESK machines.
The co-domain of continuation stores @Cont@ is comprised of 
a frame @Frame@ and a continuation address @KAddr@.
To mimic the runtime call stack, @KAddr@ plays the role of
representing the remaining stack frames.
TODO: talk about nondet of cont deref?

\begin{lstlisting}
abstract class Storable 
case class Clos(v: Lam, env: Env) extends Storable

case class Frame(x: String, e: Expr, env: Env)
case class Cont(frame: Frame, kaddr: KAddr)
\end{lstlisting}

As a consequence, the components of a state is also changed: 
the store is divided into binding store and continuation store;
the continuation stack becomes an address.
By dereferencing this address in continuation store, we can retrieve the 
actual transfer of controls. The definition of environment @Env@
remains the same.

\begin{lstlisting}
case class State(e: Expr, env: Env, bstore: BStore, 
                 kstore: KStore, k: KAddr, time: Time)
\end{lstlisting}

\subsubsection{Allocating Addresses}
Up to now, we did not describe much about the adresses in stores and the 
time-stamp @Time@.
But they are key ingredients to achieve analyses with different sensitivities,
as well as a finite state space analysis.
To effectively approximate the runtime behavior, a finite 
program contour @time@ that encodes the program execution history is introduced.
Here, we use a list of execution contexts (expressions) to reprent that, and
as we will see in section \ref{kcfainst}, by using different @tick@
function on time we may obtain a family of analyses.

\begin{lstlisting}
type Time = List[Expr]
\end{lstlisting}

As we mentioned previously, the space of addresses is finite in AAM.
Binding addresses are parameterized by variable names and the time that creates this binding,
which are both finite.
Continuation addresses @KAddr@ has two variants: 
1) @Halt@ which corresponds to the empty stack, and
2) @ContAddr@ consists of the entry expression of callee and the time that
creates this address, and they are also finite. 
(TODO: why target expression, allocation polyvariance)

\begin{lstlisting}
case class BAddr(x: String, time: Time)

abstract class KAddr
case object Halt extends KAddr
case class ContAddr(tgt: Expr, time: Time) extends KAddr
\end{lstlisting}

We use two helper functions @allocBind@ and @allocKont@ that will be used later
to allocate binding addresses and continuation addresses.

\begin{lstlisting}
def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
def allocKont(tgtExpr: Expr, time: Time): KAddr = ContAddr(tgtExpr, time)
\end{lstlisting}

@@

Since dereferencing an address becomes non-deterministic, our @atomicEval@
function also truns into non-deterministic. Given an atomic expression @e@,
@atomicEval@ returns a set of storable values, i.e., closures, to the caller.
If the expression is just a @lambda@ term, then the returned set is singleton.

\begin{lstlisting}
def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
  case Var(x) => bstore(env(x))
  case lam@Lam(x, body) => Set(Clos(lam, env))
}
\end{lstlisting}

The structure of function @step@ is similar with concrete CESK machines, 
except non-determinism which makes @step@ returns a list of reachable 
successor states.
In the beginning of @step@, we call function @tick@ to update 
the time.
Then we have two cases to consider: 1) Current control expression is @Let@,
and @App(f, ae)@ is being bind to @x@.
Fristly we retrieve the set of closures that @f@ may represents.
For each closure in the set, we do almost the same as what we did in concrete 
CESK machines. But note that since the continuation is allocated on store (@kstore@),
a new continuation address @new_kaddr@ is constructed, and a new frame
@Frame(x, e, env)@ paired with the current continuation address @kaddr@
is joined into this address. Finally, a list of states is generated.

2) An atomic expression @ae@ sits on the control position of the state.
The value of @ae@ is being returned to its caller. 
We dereference the continuation address @kaddr@ and obtain a set of
continuations @conts@.
For each continuation in the set, we construct an environment based on the
environment of the frame (@env_f@), and bind @x@ to a newly created binding 
address @baddr@; also update the store with @baddr@ and 
the values that @ae@ represents. In every generated states,
the control becomes the expression @e@ in the frame,
and as we can tell from the name, continuation address @f_kaddr@
also comes from the frame.

\begin{lstlisting}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) 
      if isAtomic(ae) =>
      val closures = atomicEval(f, env, bstore).toList
      for (Clos(Lam(v, body), env_c) <- closures) yield {
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
        val new_kaddr = allocKont(body, new_time)
        val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
        State(body, new_env, new_bstore, new_kstore, new_kaddr, new_time)
      }
    case State(ae, env, bstore, kstore, kaddr, time) 
      if isAtomic(ae) =>
      val conts = kstore(kaddr).toList
      for (Cont(Frame(x, e, env_f), f_kaddr) <- conts) yield {
        val baddr = allocBind(x, new_time)
        val new_env = env_f + (x -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(e, new_env, new_store, kstore, f_kaddr, new_time)
      }
  }
}
\end{lstlisting}

@@

TODO

\begin{lstlisting}
def k: Int = 0
def tick(s: State): Time = (s.e :: s.time).take(k)
\end{lstlisting}

@@

To run (analyze) a program, as CESK machine, we use @inject@ to construct
the initial state given a program. 
Note that the initial continuation store has a special mapping that maps continuation address
@Halt@ to empty set of continuations.
We also provide an empty program contour as initial time.
\begin{lstlisting}
def inject(e: Expr): State = 
  State(e, Map(), 
        Store[BAddr, Storable](Map()), 
        Store[KAddr, Cont](Map(Halt -> Set())), 
        Halt, List())
\end{lstlisting}

However, comparing with concrete CESK machine, the
@drive@ function performs collecting semantics instead of valuation
semantics. That is, for the purpose of analyzing programs, the function @drive@
collects all the intermediate machine states as the program is abstractly executing.
The following code shows a variant of work list algorithm that always applies @step@
function to the first element @hd@ of the work list @todo@ if @hd@
is unseen, then adds the result of @step@ to the rest of work list and in 
the meanwhile adds @hd@ to the explored states set.
If the work list is empty, the function just returns the set of reachable states up to now.

\begin{lstlisting}
def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
  case Nil => seen
  case hd::tl if seen.contains(hd) => drive(tl, seen)
  case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
}

def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
\end{lstlisting}

Finally, a user could use function @analyze@ to obtain all reachable states 
for a given program.

@@

\begin{lstlisting}
(let ([id (lambda (z) z)])
  (let ([x (id 1)])
    (let ([y (id 2)])
      x)))
\end{lstlisting}

In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
the return-flow of invocation @(id 2)@
is going to both call-sites @(id 1)@ and @(id 1)@. Therefore, the
returned value @2@ for variable @y@ is also propagated to 
variable @x@ and then imprecise analysis result is arose.
This return-flow merging is inevitable even we increase the context-sensitivity.
If we use the monovariant analysis 0CFA, the analyzing result would be @x@ and 
@y@ point to set @\{1, 2\}@, because it not distinguishes
@z@ at different call-sites.
Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
@id@ has two different values at two call-sites, so variable @y@ 
would not be polluted by @1@, but it is still helplessness for distinguish
return-flows.
TODO: do need more details to explain this example? (as in P4F paper)

@@

In the definition of state @State@, continuation store disappears, and
@konts@ becomes a list of frames. Other components keep the same.

\begin{lstlisting}
case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
\end{lstlisting}

The state transition function @step@ is still non-determinism, but
the only non-determinism happens when dereferencing the callee @f@ 
from the store at call-site @App(f, ae)@. 
For each closure of @f@ in the set, we do exactly
the same as what we did in concrete CESK machines: a new frame is constructed
and pushed onto the stack.
In the second case, since the continuation @konts@ is a list of 
frames, we can simply decompose it to differentiate an empty stack
or a frame exists on the top of stack. 
If the stack is empty, then empty successors is returned, because the computation
is finished.
Otherwise, we matched a frame @Frame(x, e, env_f)@. 
This transition is deterministic, because we only have one frame, this frame
is then popped and the new state uses the rest of stack.

\begin{lstlisting}
def step(s: State): List[State] = {
  val new_time = tick(s)
  s match {
    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
        val frame = Frame(x, e, env)
        val baddr = allocBind(v, new_time)
        val new_env = env_c + (v -> baddr)
        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
        State(body, new_env, new_store, frame::konts, new_time)
      }
    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
      konts match {
        case Nil => List()
        case Frame(x, e, env_f)::konts =>
          val baddr = allocBind(x, new_time)
          val new_env = env_f + (x -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          List(State(e, new_env, new_store, konts, new_time))
      }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
(letrec ([f1 (lambda (x) 
               (let ([x1 (f2 x)]) x1))]
         [f2 (lambda (y)
               (let ([y1 (f1 y)]) y1))])
  (let ([z (f1 1)])
    z))
\end{lstlisting}

@f1@ and @f2@ call each others, and the stack will be interleaved
pushing @f1@ and @f2@ onto the top of current stack, but not existing 
two stack components are identical in the state space.

@@

\begin{lstlisting}
case class P4FContAddr(tgt: Expr, tgtEnv: Env, time: Time) extends KAddr
def allocKont(tgtExpr: Expr, tgtEnv: Env, time: Time): KAddr = 
  P4FContAddr(tgtExpr, tgtEnv, time)
\end{lstlisting}

@@

\begin{lstlisting}
def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
  case Nil => seen
  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
  case hd::tl =>
    val new_time = tick(hd)
    val new_states = hd match {
      case State(Let(x, App(f, ae), e), env, bstore, konts, time) 
        if isAtomic(ae) =>
        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
          val frame = Frame(x, e, env)
          val baddr = allocBind(v, new_time)
          val new_env = env_c + (v -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          State(body, new_env, new_store, frame::konts, new_time)
        }
      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
        konts match {
          case Nil => List()
          case Frame(x, e, f_env)::konts =>
            val baddr = allocBind(x, new_time)
            val new_env = f_env + (x -> baddr)
            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
            List(State(e, new_env, new_store, konts, new_time))
        }
    }
    drive_step(new_states ++ tl, seen + hd)
}

def analyze(e: Expr): Set[State] = drive_step(List(inject(e)), Set())
\end{lstlisting}
